---
title: "Assignment_1"
author: "Krishna Kumar Tavva - 811283461"
date: "1/29/2023"
output: pdf_document
---

# Downloaded the Pharmaceuticals dataset from Canvas and stored to this PC.

# Imported the Pharmaceuticals into R

```{r}
#install.packages("tinytex")
#tinytex::install_tinytex()
```

```{r}
#install.packages("readr")
pharmaceuticals=read.csv("E:\\Fundamentals of Machine Learning\\Datasets\\Pharmaceuticals.csv")
```

# Print out descriptive statistics for a selection of quantitative and categorical variables

```{r}
summary(pharmaceuticals)

```
# Transform at least one variable

```{r}
#install.packages("tidyverse")

library(tidyverse)
```

# Logarithmic Transformation of PE_Ratio

```{r}
log10(pharmaceuticals$PE_Ratio)
```

# Histogram Plot of Asset_Turnover
 ```{r}
hist(pharmaceuticals$Asset_Turnover)
```
# Scatter Plot between ROE & ROA

```{r}
plot(pharmaceuticals$ROE, pharmaceuticals$ROA)
```

# Catregorical Plot for Median Recommendation

```{r}
#install.packages("ggplot2")
library(ggplot2)
ggplot(pharmaceuticals, aes(x=pharmaceuticals$Median_Recommendation)) +geom_bar(fill='dark grey') +  labs(x='MR')
```




